Data Exploration

### 2018-11-14 src/data_exploration.Rmd
remvoing stuff ive already done (doh)
simple plots of env. correlates

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
```{r pre-install packages, message=FALSE}
packages <- c() 
source("../src/ipak.R")
ipak(packages)
```

@@

>>>>

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
The "include=FALSE" supresses the package installation text appearing in the document...
```{r pre-install packages, include=FALSE}
packages <- c("plyr", "graphics") 
source("../src/ipak.R")
ipak(packages)
```

@@

# environmental correlates

1) Get the max, min, an mean values and add into a dataframe

Temperature
```{r}
temp_depth_max <- max(presence$temp_depth, na.rm = TRUE)
temp_depth_min <- min(presence$temp_depth, na.rm = TRUE)
temp_depth_mean <- mean(presence$temp_depth, na.rm = TRUE)
temp_surface_max <- max(presence$temp_surface, na.rm = TRUE)
temp_surface_min <- min(presence$temp_surface, na.rm = TRUE)
temp_surface_mean <- mean(presence$temp_surface, na.rm = TRUE)
```

Salinity
```{r}
sal_depth_max <- max(presence$salinity_depth, na.rm = TRUE)
sal_depth_min <- min(presence$salinity_depth, na.rm = TRUE)
sal_depth_mean <- mean(presence$salinity_depth, na.rm = TRUE)
sal_surface_max <- max(presence$salinity_surface, na.rm = TRUE)
sal_surface_min <- min(presence$salinity_surface, na.rm = TRUE)
sal_surface_mean <- mean(presence$salinity_surface, na.rm = TRUE)
```

Chl
```{r}
chl_depth_max <- max(presence$chl_depth, na.rm = TRUE)
chl_depth_min <- min(presence$chl_depth, na.rm = TRUE)
chl_depth_mean <- mean(presence$chl_depth, na.rm = TRUE)
chl_surface_max <- max(presence$chl_surface, na.rm = TRUE)
chl_surface_min <- min(presence$chl_surface, na.rm = TRUE)
chl_surface_mean <- mean(presence$chl_surface, na.rm = TRUE)
```

O2
```{r}
o2_depth_max <- max(presence$o2_depth, na.rm = TRUE)
o2_depth_min <- min(presence$o2_depth, na.rm = TRUE)
o2_depth_mean <- mean(presence$o2_depth, na.rm = TRUE)
o2_surface_max <- max(presence$o2_surface, na.rm = TRUE)
o2_surface_min <- min(presence$o2_surface, na.rm = TRUE)
o2_surface_mean <- mean(presence$o2_surface, na.rm = TRUE)
```

MLP
```{r}
mlp_surface_max <- max(presence$mlp_surface, na.rm = TRUE)
mlp_surface_min <- min(presence$mlp_surface, na.rm = TRUE)
mlp_surface_mean <- mean(presence$mlp_surface, na.rm = TRUE)
```

SSH
```{r}
ssh_surface_max <- max(presence$ssh_surface, na.rm = TRUE)
ssh_surface_min <- min(presence$ssh_surface, na.rm = TRUE)
ssh_surface_mean <- mean(presence$ssh_surface, na.rm = TRUE)
```

create matrix
```{r}
td <- c(temp_depth_max, temp_depth_min, temp_depth_mean)
ts <- c(temp_surface_max, temp_surface_min, temp_surface_mean)
sd <- c(sal_depth_max, sal_depth_min, sal_depth_mean)
ss <- c(sal_surface_max, sal_surface_min, sal_surface_mean)
cd <- c(sal_depth_max, sal_depth_min, sal_depth_mean)
cs <- c(sal_surface_max, sal_surface_min, sal_surface_mean)
od <- c(o2_depth_max, o2_depth_min, o2_depth_mean)
os <- c(o2_surface_max, o2_surface_min, o2_surface_mean)
mlp <- c(mlp_surface_max, mlp_surface_min, mlp_surface_mean)
ssh <- c(ssh_surface_max, ssh_surface_min, ssh_surface_mean)
env_stats <- rbind(td, ts, sd, ss, cd, cs, od, os, mlp, ssh)
row.names(env_stats) <- c("Temp Depth", "Temp Surface", "Salinity Depth", "Salinity Surface", "Chl Depth", "Chl Surface", "Oxygen Depth", "Oxygen Surface", "MLP", "SSH") 
colnames(env_stats) <- c("Max", "Min", "Mean")
write.csv(env_stats, "../output/env/env_correlates_basic_stats.csv", row.names = TRUE)
env_stats
```

Hmm some potentially suspicious values here in
- Oxygen Depth min
- mlp  Max
and maybe also
- Salinity Depth min
- Salinity Surface min
- CHL depth min
- CHL depth max

lets just see where these values appear and check the netcdf layers. Note I will check these values in cygwin using the cdo operator infov on the netcdfs (to ensure there was no issue with processing in R).

get the year and month these values appeared in
```{r}
o2d_check_yr <- subset(presence$year, presence$o2_depth == o2_depth_min)
o2d_check_mth <- subset(presence$month, presence$o2_depth == o2_depth_min)
```
O2 depth min is in 2005_08

And value seems correct

```{r}
mlp_check_yr <- subset(presence$year, presence$mlp_surface == mlp_surface_max)
mlp_check_mth <- subset(presence$month, presence$mlp_surface == mlp_surface_max)
```
mlp max in in 2007_12

And yes the value seems correct

```{r}
sald_check_yr <- subset(presence$year, presence$salinity_depth == sal_depth_max)
sald_check_mth <- subset(presence$month, presence$salinity_depth == sal_depth_max)
```
year month 2001_05

also seems correct

```{r}
sals_check_yr <- subset(presence$year, presence$salinity_surface == sal_surface_min)
sals_check_mth <- subset(presence$month, presence$salinity_surface == sal_surface_min)
```
year month 21998_06

ok again...

Leave the rest

#simple plots of env. correlates

temperature
```{r}
par(mfrow=c(1,2))
plot(presence$temp_depth, main = "Temp at Depth (Various)", ylab = "Kelvin")
plot(presence$temp_surface, main = "Temp at Surface", ylab = "Kelvin")
dev.copy(png, "../output/env/simple_plots/temperature_simpleplot.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```
Hmm not necessarily very helpful. But anyway, lets carry on

salinity
```{r}
par(mfrow=c(1,2))
plot(presence$salinity_depth, main = "Salinity at Depth (Various)", ylab = "Practical Salinity Units")
plot(presence$salinity_surface, main = "Salinity at Surface", ylab = "Practical Salinity Units")
dev.copy(png,"../output/env/simple_plots/salinity_simpleplot.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

Chlorophyll
```{r}
par(mfrow=c(1,2))
plot(presence$chl_depth, main = "Chl at Depth (Various)", ylab = "Chl (mmol.m-3)")
plot(presence$chl_surface, main = "Chl at Surface", ylab = "Chl (mmol.m-3)")
dev.copy(png,"../output/env/simple_plots/chl_simpleplot.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

O2
```{r}
par(mfrow=c(1,2))
plot(presence$o2_depth, main = "Dissolved Oxygen at Depth (Various)", ylab = "Chl (mmol.m-3)")
plot(presence$o2_surface, main = "Dissolved Oxygen at Surface", ylab = "O2 (mmol.m-3)")
dev.copy(png,"../output/env/simple_plots/o2_simpleplot.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

mlp
```{r}
plot(presence$mlp_surface, main = "Mixed Layer Thickness", ylab = "Depth (m)")
dev.copy(png,"../output/env/simple_plots/mlp_simpleplot.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

SSH
```{r}
plot(presence$ssh_surface, main = "Sea Surface Height", ylab = "Height (m)")
dev.copy(png,"../output/env/simple_plots/ssh_simpleplot.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```
